### 2011-11-08 supplement_history.tex
Added suppl history document. Updated refs

\documentclass{sig-alternate}

\begin{document}

\title{Supplement History of Chem(o)informatics as part of Cheminformatics: The Computer Science of Chemical Discovery}
\numberofauthors{9}
\author{
\alignauthor
Joerg Kurt Wegner\\
       \affaddr{Tibotec BVBA}\\
       \affaddr{Turnhoutseweg 30}\\
       \affaddr{2340 Beerse Turnhout, Belgium}\\
       \email{jwegner@its.jnj.com}
% 2nd. author
\alignauthor
Aaron Sterling\\
       \affaddr{Department of Computer Science}\\
       \affaddr{Iowa State University}\\
       \affaddr{Ames, Iowa, USA}\\
       \email{sterling@iastate.edu}
% 3rd author
\alignauthor
Rajarshi Guha\\
\affaddr{NIH Center for Translational Therapeutics}\\
\affaddr{9800 Medical Center Drive}\\
\affaddr{Rockville, MD 20850}\\
\email{guhar@mail.nih.gov}
}

\additionalauthors{Additional authors:
Andreas Bender (University of Cambridge, email: {\texttt{andreas.bender@cantab.net}}),
Jean-Loup Faulon (University of Evry, email: {\texttt{Jean-Loup.Faulon@issb.genopole.fr}}),
Janna Hastings (European Bioinformatics Institute, Cambridge, UK, email: {\texttt{janna.hastings@gmail.com}}),
Noel O'Boyle (University College Cork, Cork, Ireland, email: {\texttt{baoilleach@gmail.com}}),
John Overington (European Bioinformatics Institute, Cambridge, UK, email: {\texttt{jpo@ebi.ac.uk}}),
Herman Van Vlijmen (Tibotec, Beerse, Belgium, email: {\texttt{hvvlijme@its.jnj.com}}), and
Egon Willighagen (Karolinska Institutet, Stockholm, Sweden, email: {\texttt{egon.willighagen@ki.se}})
.}
\date{25 June 2011}
\maketitle

\section{History of Chem(o)informatics}
The aim of this brief historic review of cheminformatics is to put the
content of the main article into a broader perspective. Therefore it
might be useful to clarify some other terms and how cheminformatics
might relate to them.  Still, Chen \cite{Chen2006} and Brown
\cite{brown2009} cited various definitions of pioniers in the field
and Brown concluded rightfully: '\textit{The differences in
  definitions (of the term cheminformatics) are largely a result of
  the types of analyses that particular scientists practice and no
  single definition is intended to be all-encompassing}.'
\begin{itemize}
\item \textbf{Quantum chemistry} is something most chemists would
  relate with 'theoretical chemistry'. In short, it allows to describe
  chemical molecules on an electron density level, which allows also
  to describe physical properties, e.g infrared spectra of a chemical
  compound.  The results of quantum chemistry calculation are used in
  molecular modeling (energy force field parameters) and
  cheminformatics (atom partial charges, or full molecule properties,
  e.g molecule polarizabilities).
\item \textbf{Molecular modeling} or \textbf{Computational chemistry}
  considers typically energy calculations of single molecules or
  interactions with other molecules, e.g. proteins. 'Molecular
  docking' belongs into this category, since it places a chemical
  molecule in the active site of enzymes and calculates
  energy/interaction scores between a chemical molecule and the
  protein. Surely are there various cheminformatics concepts required
  to perform such a task. Still, strictly speaking is cheminformatics
  rather on the chemical molecule side than on the protein side,
  though are both sides affecting each other. This area is typically
  strongly related to structural biology and technologies like XRay
  crystallography and NMR spectra structure determination.
\item \textbf{Chemometrics} can be considered as applied
  cheminformatics to the field of chemical/physical analysis,
  e.g. infrared, mass and NMR spectra analysis.
\item '\textbf{Translational research} aims to improve the health and
  longevity of the world's populations and depends on developing
  broad-based teams of scientists and scholars who are able to focus
  their efforts to link basic scientific discoveries with the arena of
  clinical investigation, and translating the results of clinical
  trials into changes in clinical practice, informed by evidence from
  the social and political sciences.' (Source:Wikipedia).  Since
  clinical trials can be connected to drugs, chemical molecules, even
  all of translational research and biobanking could be considered as
  a field of cheminformatics, which would be on the other hand just a
  small contribution to a full knowledge management in this field
  \cite{Szalma2010}.
\end{itemize}
If we consider all information, analysis, and optimization of a
chemical molecule as 'cheminformatics', then the field is very large,
since a chemical molecule plays the central role in many related
disciplines. Still, for keeping it practical the aim of this review is
not being exhaustive, but rather to point the reader to more detailed
references and to highlight some historic milestones.

One of the most comprehensive cheminformatic series was published by
Prof. Gasteiger in 2003 \cite{Gasteiger2003}, one of the pioniers in
the field.  Compacter introductory books followed from Gasteiger/Engel
\cite{gasteigerengel2003} and Leach/Gillet \cite{leachgillet2007}.
The books of Bajorath \cite{Bajorath2004} and Oprea \cite{oprea2005}
focus on the applications of cheminformatics in drug design.  The
first books discussing chemical graph theory and algorithms were
published in 1989 by Zupan \cite{zupan1989}, 1991 by Bonchev/Rouvrey
\cite{bonchevrouvrey1991,bonchevrouvrey2003} and 1992 by Trinastic
\cite{Trinajstic1992}.  The most recent one was published by
Faulon/Bender \cite{faulon2010}.  Other books focus more on specific
fields/applications like mathematical challenges
\cite{mathchallenges1995}, molecular diversity
\cite{moleculardiversity1999}, factor analysis \cite{Malinowski2002},
evolutionary algorithms \cite{clark2000}, molecular descriptors
\cite{todeschini2000}.

\section{Historic milestones}
The first chemical graphs were drawn by the Scottish chemist Willam
Cullen in 1758 \cite{bonchevrouvrey1991}, which called them initially
affinity graphs.  After developing the concept of bonds between atoms
(Couper, 1858) the first chemical graphs occured in publications of
Brown in 1864 and Cayley in 1874 \cite{bonchevrouvrey1991,brown2009}.

The year 1946 may be regarded as the birth year of chemoinformatics
\cite{Chen2006}: '\textit{In 1946 King et al.\cite{kct1946} published
  an article illustrating the use of IBM's business accounting
  machines in carrying out the construction of the rotational spectra
  of asymmetric rotors by the evaluation of mathematical equations for
  line position and line intensity.}'

The first records of managing chemical information go back to chemical
literature being indexed since the year 1771.  In 1881 the first
edition of \textit{Beilstein's Handbuch der Organischen Chemie}
ecyclopedia was published \cite{polanski2009} registering 1500
chemical compounds. The short term 'chemoinformatics' was coined by
Brown in 1998\cite{brown1998}, while the \textit{Journal of Chemical
  Documentation} (founded in 1961) changed already its name in 1975 to
\textit{Journal of Chemical Information and Computer Sciences} to
reflect the tight connection between chemical information and computer
science. The Chemical Abstract Service (CAS) of the American Chemical
Society (ACS) formed in 1955 a research department and starting from
1965 they provided their Chemical Registry System, and in 1968 they
made the first computer-readable file of all abstracted documents
available \cite{Chen2006}.

Ray and Kirsch published in 1957 the first substructure searching
algorithm for allowing to find chemical records via computers
\cite{RayKirsch1957}.

One of the key algorithms for molecular graph canonicalization is the
Morgan algorithm \cite{Morgan1965}. It will search for the most
central bond in a graph and allows then to traverse other atoms,
resolve ties, and assigning new atom indices. Still, there are various
options how different software implementations are handling the 'ties'
for assigning new atom indices. So, by using the canonicalization of
multiple software implementations and chemical expert systems can be
even used to create randomized (canonical, only when sticking to one
software) molecular graphs, which can be used as stochastic input for
other algorithms \cite{cokfl06}. The central bond is also being used
to create 3D conformational search trees \cite{Schwab2001}.

In 1965 the DENDRAL expert system started with the aim to
automatically determine the structure of an unknown chemical compound
from the corresponding mass spectrum \cite{Gray1986}. The system is
also often being cited as one of the earliest artificial intelligence
and expert systems \cite{Chen2006}.

The first system supporting chemical synthesis planning was OCSS
(Organic Chemical Simulation of Syntheses) developed by Corey and
Wipke in 1969 \cite{CoreyWipke1969}.

In the early 1970s the collaboration between Lempel-Ziv-Welch has led
to the LZW algorithm, since the Cheminformatician Lynch was discussing
chemical fragment screening approaches with the computer scientists
Lempel and Ziv \cite{brown2009}.

In 1978 Gasteiger/Marsili published a fast algorithm to calculate
partial charges in organic molecules by a Partial Equalization of
Orbital Electronegativities (PEOE) \cite{gm78}, which is till today
one of the gold-standards to calculate partial charges of atoms.

Instead of converting molecular graphs to a vector representation for
applying machine learning methods, it is also possible to use direct
molecular graph mining methods \cite{okada2006}. Some of the first
published methods use neural networks \cite{kireev1995}, inductive
logic programming \cite{yh02a}, or graph kernels \cite{kti03}.

Molecular docking of chemical molecules into proteins is an important
application in molecular modeling \cite{Leach2001}. It is noteworthy
that in some cases molecular shape matching methods alone
(cheminformatics), can outperform energy-based molecular docking
algorithms (molecular modeling) \cite{hsn07}.

\bibliographystyle{abbrv}
\bibliography{paper}

\end{document}