### 2011-11-09 supplement_history.tex
more updates to suppl history

\documentclass{sig-alternate}

\begin{document}

\title{Supplement History of Cheminformatics as part of Cheminformatics: The Computer Science of Chemical Discovery}
\numberofauthors{9}
\author{
\alignauthor
Joerg Kurt Wegner\\
       \affaddr{Tibotec BVBA}\\
       \affaddr{Turnhoutseweg 30}\\
       \affaddr{2340 Beerse Turnhout, Belgium}\\
       \email{jwegner@its.jnj.com}
% 2nd. author
\alignauthor
Aaron Sterling\\
       \affaddr{Department of Computer Science}\\
       \affaddr{Iowa State University}\\
       \affaddr{Ames, Iowa, USA}\\
       \email{sterling@iastate.edu}
% 3rd author
\alignauthor
Rajarshi Guha\\
\affaddr{NIH Center for Translational Therapeutics}\\
\affaddr{9800 Medical Center Drive}\\
\affaddr{Rockville, MD 20850}\\
\email{guhar@mail.nih.gov}
}

\additionalauthors{Additional authors:
Andreas Bender (University of Cambridge, email: {\texttt{andreas.bender@cantab.net}}),
Jean-Loup Faulon (University of Evry, email: {\texttt{Jean-Loup.Faulon@issb.genopole.fr}}),
Janna Hastings (European Bioinformatics Institute, Cambridge, UK, email: {\texttt{janna.hastings@gmail.com}}),
Noel O'Boyle (University College Cork, Cork, Ireland, email: {\texttt{baoilleach@gmail.com}}),
John Overington (European Bioinformatics Institute, Cambridge, UK, email: {\texttt{jpo@ebi.ac.uk}}),
Herman Van Vlijmen (Tibotec, Beerse, Belgium, email: {\texttt{hvvlijme@its.jnj.com}}), and
Egon Willighagen (Karolinska Institutet, Stockholm, Sweden, email: {\texttt{egon.willighagen@ki.se}})
.}
\date{25 June 2011}
\maketitle

\section{History of Cheminformatics}
The aim of this brief review of the history of cheminformatics is to
put the content of the main article into a broader perspective. In
addition, we also discuss some terms as they relate to
cheminformatics.  It is thus useful to first provide some definitions
of cheminformatics itself, given that they also shed some light on the
founding principles of the field. For example, we have the following
definitions:
\begin{itemize}
\item ``\textit{The mixing of information resources to transform data
    into information, and information into knowledge, for the intended
    purpose of making better decisions faster in the arena of drug
    lead identification and optimization.}'' Frank K. Brown, 1998.
\item ``\textit{[Chemoinformatics involves]... the computer
    manipulation of two- or threedimensional chemical structures and
    excludes textual information. This distinguishes the term from
    chemical information, largely a discipline of chemical librarians
    and does not include the development of computational methods.}''
  Peter Willett, 2002.
\item ``\textit{\ldots the application of informatics to solve chemical
    problems.}'' and ``\textit{\ldots chemoinformatics makes the point that
    you're using one scientific discipline to understand another
    scientific discipline.}'' Johann Gasteiger, 2002
\end{itemize}
More generally, Chen\cite{Chen2006} and Brown \cite{brown2009} cited various
definitions of pioneers in the field with Brown concluding that
``'\textit{The differences in definitions (of the term
  cheminformatics) are largely a result of the types of analyses that
  particular scientists practice and no single definition is intended
  to be all-encompassing}'', as is evident from the three definitions above.

It is also useful to clarify some other terms and how
cheminformatics might relate to them:
\begin{itemize}
\item \textbf{Quantum chemistry} (QM) is generally associated with
  theoretical chemistry or chemical physics. Briefly, it focuses on
  the description of chemical systems starting from first principles
  and is based on the Schr\"{o}dinger wave equation and frameworks
  built on top of it. The use of quantum mechanics allows one to
  describe molecules in terms of electron density, and hence calculate
  molecular properties (spectra, energies et.) in a physically
  accurate manner.  The results of quantum chemistry calculations are
  used in molecular modeling (energy force field parameters) and
  cheminformatics (atom partial charges, or full molecule properties,
  e.g molecule polarizabilities).
\item \textbf{Molecular modeling} or \textbf{Computational chemistry}
  can be considered an approximation of quantum chemistry and aims to
  evaluate many of the properties considered by QM methods. The key
  difference is that these methods employ a variety of
  parametrizations (usually, but not always, derived from QM studies),
  allowing one to evaluate energies for small molecules very rapidly
  as well as handle larger systems (such as proteins) that are too
  time consuming for QM methods. A variety of techniques including
  semi-empirical methods, molecular mechanics and docking belong to
  this category. While it is true that many of these methods involve
  cheminformatics concepts (choice of tautomer, partial charge
  assignments, etc.), the methodologies are fundamentally physical in
  nature. We note that computational chemistry has close ties to
  topics such as x-ray crystallography and NMR structure elucidation.
\item \textbf{Chemometrics} is a focused application of statistical
  methods to problems in analytical chemistry, such as infrared, mass
  and NMR spectroscopy.
\item \textbf{Translational research/medicine} aims to make more
  direct links between academic (or early stage) research and clinical
  practice in an effort to speed up the results of research into
  actual treatments. When applied to medical research this usually
  implies efforts to speed up the process by which a compounds goes
  from the lead stage ot the clinical trial stage. Naturally,
  cheminformatics plays a role at various stages of this process (lead
  identification, lead optimization, ADMET modeling) as well as more
  broadly such as in drug repurposing efforts.
\end{itemize}
If we consider all information, analysis, and optimization of a
molecule as ``cheminformatics'', then the field is very large, since a
molecule plays the central role in many related disciplines. Clearly,
this article is not mean to be an exhaustive overview; rather, we
point the reader to more detailed references and to highlight some
historic milestones.

One of the most comprehensive cheminformatic series was published by
Prof. Gasteiger in 2003 \cite{Gasteiger2003}, one of the pioniers in
the field.  Compacter introductory books followed from Gasteiger/Engel
\cite{gasteigerengel2003} and Leach/Gillet \cite{leachgillet2007}.
The books of Bajorath \cite{Bajorath2004} and Oprea \cite{oprea2005}
focus on the applications of cheminformatics in drug design.  The
first books discussing chemical graph theory and algorithms were
published in 1989 by Zupan \cite{zupan1989}, 1991 by Bonchev/Rouvrey
\cite{bonchevrouvrey1991,bonchevrouvrey2003} and 1992 by Trinastic
\cite{Trinajstic1992}.  The most recent one was published by
Faulon/Bender \cite{faulon2010}.  Other books focus more on specific
fields/applications like mathematical challenges
\cite{mathchallenges1995}, molecular diversity
\cite{moleculardiversity1999}, factor analysis \cite{Malinowski2002},
evolutionary algorithms \cite{clark2000}, molecular descriptors
\cite{todeschini2000}.

\section{Historic milestones}
The first chemical graphs were drawn by the Scottish chemist Willam Cullen in 1758 \cite{bonchevrouvrey1991}, which called them initially affinity graphs.
After developing the concept of bonds between atoms (Couper, 1858) the first chemical graphs occured in publications 
of Brown in 1864 and Cayley in 1874 \cite{bonchevrouvrey1991,brown2009}.

The year 1946 may be regarded as the birth year of chemoinformatics \cite{Chen2006}: '\textit{In 1946 King et al.\cite{kct1946} published an article illustrating 
the use of IBM's business accounting machines in carrying out the construction of the rotational spectra of asymmetric rotors
by the evaluation of mathematical equations for line position and line intensity.}'

The first records of managing chemical information go back to chemical literature being indexed since the year 1771.
In 1881 the first edition of \textit{Beilstein's Handbuch der Organischen Chemie} ecyclopedia was 
published \cite{polanski2009} registering 1500 chemical compounds. The short term 'chemoinformatics' was coined by Brown 
in 1998\cite{brown1998}, while the \textit{Journal of Chemical Documentation} (founded in 1961) changed already its name in 1975 to 
\textit{Journal of Chemical Information and Computer Sciences} to reflect the tight connection between chemical information 
and computer science. The Chemical Abstract Service (CAS) of the American Chemical Society (ACS) formed in 1955 a research department
and starting from 1965 they provided their Chemical Registry System, and in 1968 they made the first computer-readable file of all
abstracted documents available \cite{Chen2006}.

Ray and Kirsch published in 1957 the first substructure searching algorithm for allowing to find chemical records via computers \cite{RayKirsch1957}.

One of the key algorithms for molecular graph canonicalization is the Morgan algorithm \cite{Morgan1965}. It will search for the most central bond
in a graph and allows then to traverse other atoms, resolve ties, and assigning new atom indices. Still, there are various options how different 
software implementations are handling the 'ties' for assigning new atom indices. So, by using the canonicalization of multiple software implementations and chemical
expert systems can be even used to create randomized (canonical, only when sticking to one software) molecular graphs, which can be used as stochastic input for 
other algorithms \cite{cokfl06}. The central bond is also being used to create 3D conformational search trees \cite{Schwab2001}.

In 1965 the DENDRAL expert system started with the aim to automatically determine the structure of an unknown chemical compound from the 
corresponding mass spectrum \cite{Gray1986}. The system is also often being cited as one of the earliest artificial intelligence and expert systems  \cite{Chen2006}.

The first system supporting chemical synthesis planning was OCSS (Organic Chemical Simulation of Syntheses) developed by Corey and Wipke in 1969 \cite{CoreyWipke1969}.

In the early 1970s the collaboration between Lempel-Ziv-Welch has led to the LZW algorithm, since the Cheminformatician Lynch was
discussing chemical fragment screening approaches with the computer scientists Lempel and Ziv \cite{brown2009}.

In 1978 Gasteiger/Marsili published a fast algorithm to calculate partial charges in organic molecules by a Partial Equalization of Orbital 
Electronegativities (PEOE) \cite{gm78}, which is till today one of the gold-standards to calculate partial charges of atoms.

Instead of converting molecular graphs to a vector representation for applying machine learning methods, it is also possible to use direct 
molecular graph mining methods \cite{okada2006}. Some of the first published methods use neural networks \cite{kireev1995}, 
inductive logic programming \cite{yh02a}, or graph kernels \cite{kti03}.

Molecular docking of chemical molecules into proteins is an important application in molecular modeling \cite{Leach2001}. It is noteworthy that in some cases
molecular shape matching methods alone (cheminformatics), can outperform energy-based molecular docking algorithms (molecular modeling) \cite{hsn07}.

\bibliographystyle{abbrv}
\bibliography{paper}

\end{document}